\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-11 k-planar.tex
Reworked opening paragraphs. We don't need to sell so hard.

\note{PM}{Agreed. Done.}

@@

The starting point for this work is the following `product structure theorem' for planar graphs\footnote{In this paper, all graphs are finite and undirected. Unless mentioned otherwise, all graphs are also simple. For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$), let $G[S]$  denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.  We use $G-S$ as a shorthand for $G[V(G)\setminus S]$. Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (improved by \citet{UWY}). A graph $G$ is \defin{contained in} a graph $X$ if $G$ is isomorphic to a subgraph of $X$.

@@

\noindent\cref{PlanarProduct} has been the key tool in solving the following well-known open problems:
\begin{compactitem}
\item \citet{DJMMUW20} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92}).
\item  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02}).
\item \citet{DFMS21} use it to make dramatic improvements to the best known bounds for $p$-centered colourings of planar graphs.
\item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results starting with the work of \citet{kannan.naor.ea:implicit}). \citet{DEJGMM21} have since used it to find asymptotically optimal adjacency labellings of planar graphs.
\item The result of \citet{DEJGMM21} implies that, for every integer
$n>0$, there is a `induced-universal' graph $U_n$ with $n^{1+o(1)}$ vertices such that every $n$-vertex planar graph is an induced
subgraph of $U_n$. This has recently been extended by \citet{EJM}, who show the existence of a `subgraph-universal' graph with $(1+\epsilon)n$ vertices that contains every $n$-vertex planar graph as a subgraph as well as the existence of an `induced universal' graph with $n^{1+o(1)}$ vertices \emph{and} edges.  The former result is the first progress on subgraph-universal graphs for planar graphs since the $O(n^{3/2})$-vertex construction of \citet{babai.chung.ea:on}.
% aking the first progress on this problem since the work of \citet{babai.xx} in 1984.
\end{compactitem}

\note{DW}{Mention other application such as vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, odd colourings~\citep{DMO}? We write a lot about adjacency labellings and universal induced subgraphs. Do we need so much detail here? We also mention this work in \cref{WrappingUp}.}

>>>>

% \note{PM}{Agreed. Done.}

@@

The starting point for this work is the following `product structure theorem' for planar graphs\footnote{In this paper, all graphs are finite and undirected. Unless mentioned otherwise, all graphs are also simple. For any graph $G$ and any set $S$ (typically $S\subseteq V(G)$), let $G[S]$  denote the graph with vertex set $V(G)\cap S$ and edge set $\{uv\in E(G) : u,v\in S\}$.  We use $G-S$ as a shorthand for $G[V(G)\setminus S]$. Undefined terms are in \citep{Diestel5}.} by \citet{DJMMUW20} (with an improvement by \citet{UWY}). A graph $G$ is \defin{contained in} a graph $X$ if $G$ is isomorphic to a subgraph of $X$.

@@

\noindent\cref{PlanarProduct} has been the key tool in solving the following well-known open problems:
\begin{compactitem}
\item \citet{DJMMUW20} use it to prove that planar graphs have bounded queue-number (resolving a conjecture of \citet{HLR92}).
\item  \citet{dujmovic.esperet.ea:planar} use it to prove that planar graphs have bounded non-repetitive chromatic number (resolving a conjecture of \citet{AGHR-RSA02}).
\item \citet{DEJGMM21} have since used it to find asymptotically optimal adjacency labellings of planar graphs (resolving a problem of \cite{kannan.naor.ea:implicit}).
\item \citet{EJM} use it to show the existence of a 'universal graph' with $n^{1+o(1)}$ vertices and edges that contains every $n$-vertex planar graph as an induced subgraph (resolving a problem of \citet{babai.chung.ea:on}).
\end{compactitem}
In addition to these rather well-known problems, \cref{PlanarProduct} has been used to resolve or make substantial progress on a number of other problems on planar graphs, including $p$-centered colouring \cite{DFMS21}, $\ell$-vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, and odd colourings~\citep{DMO}.

% \item \citet{DFMS21} use it to make dramatic improvements to the best known bounds for $p$-centered colourings of planar graphs.
% \item \citet{bonamy.gavoille.ea:shorter} use it to find shorter adjacency labellings of planar graphs (improving on a sequence of results starting with the work of \citet{kannan.naor.ea:implicit}).
% \item \citet{BDJM} use it to show that every $n$-vertex planar graph has local vertex ranking using $O(\log n/\log\log\log n)$ colours.
% % \item The result of \citet{DEJGMM21} implies that, for every integer
% $n>0$, there is a `induced-universal' graph $U_n$ with $n^{1+o(1)}$ vertices such that every $n$-vertex planar graph is an induced
% subgraph of $U_n$. This has recently been extended by \citet{EJM}, who show the existence of a `subgraph-universal' graph with $(1+\epsilon)n$ vertices that contains every $n$-vertex planar graph as a subgraph as well as the existence of an `induced universal' graph with $n^{1+o(1)}$ vertices \emph{and} edges.  The former result is the first progress on subgraph-universal graphs for planar graphs since the $O(n^{3/2})$-vertex construction of \citet{babai.chung.ea:on}.
% aking the first progress on this problem since the work of \citet{babai.xx} in 1984.
% \end{compactitem}

% \note{DW}{Mention other application such as vertex ranking~\citep{BDJM}, twin-width~\citep{BKW}, odd colourings~\citep{DMO}? We write a lot about adjacency labellings and universal induced subgraphs. Do we need so much detail here? We also mention this work in \cref{WrappingUp}.}
